# Data submission instructions

This page is intended to provide teams with all the information they need to
submit forecasts.
All forecasts should be submitted directly to the [data-processed/](./) folder.
Data in this directory should be added to the repository through a pull request
so that automatic data validation checks are run.

These instructions provide detail about the 
[data format](#Data-formatting) as well as 
[validation](#Data-validation)
that you can do prior to this pull request. 
In addition, we describe 
[meta-data](#Meta-data) that each model should provide.


*Table of Contents*

- [ground truth data](#ground-truth-data)
- [data formatting](#Data-formatting)
- [data validation](#Data-validation)
- [metadata format](#Meta-data)


## Ground truth data

There are several different sources for death data. Currently, 
all forecasts will be compared to the 
[daily reports containing death data from the JHU CSSE group](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv) 
as the gold standard reference data for deaths in the US. 
Note that there are significant differences 
(especially in daily incident death data) 
between the JHU data and another commonly used source, from the New York Times. 
The team at UTexas-Austin is tracking this issue on [a separate GitHub repository](https://github.com/spencerwoody/covid19-data-comparsion).

Weekly incident data are the sum of daily incident data from Sunday through 
Saturday. 
Weekly cumulative data is the cumulative data up to and including Saturday. 

We may add additional sources of ground-truth data at a future time.


## Data formatting

The automatic check validates both the filename and file contents to ensure the 
file can be used in the visualization and ensemble forecasting.


### Subdirectory

Each subdirectory within the [data-processed/](data-processed/) directory has 
the format

    team-model
    
where 

- `team` is the teamname and 
- `model` is the name of your model. 

Both team and model should be less than 15 characters and not include hyphens.

Within each subdirectory, there should be a metadata file, a license file
(optional), and a set of forecasts. 


### Metadata 

The metadata file should have the following format

    metadata-team-model.txt
    
and here is [the structure of the metadata file](https://github.com/reichlab/covid19-forecast-hub/blob/master/data-processed/METADATA.md).

### License (optional)

If you would like to include a license file, 
please use the following format

    LICENSE.txt

If you are not using one of the [standard licenses](https://github.com/reichlab/covid19-forecast-hub/blob/master/code/validation/accepted-licenses.csv),
then you must include a license file.



### Forecasts

Each forecast file within the subdirectory should have the following format

    YYYY-MM-DD-team-model.csv
    
where

- `YYYY` is the 4 digit year, 
- `MM` is the 2 digit month,
- `DD` is the 2 digit day,
- `team` is the teamname, and
- `model` is the name of your model. 

The date YYYY-MM-DD is the [`forecast_date`](#forecast_date).

The `team` and `model` in this file must match the `team` and `model` in the
directory this file is in. 
Both `team` and `model` should be less than 15 characters, 
alpha-numeric and underscores only, with no spaces or hyphens.


## Forecast file format

The file must be a comma-separated value (csv) file with the following 
columns (in any order):

- `forecast_date`
- `target`
- `target_end_date`
- `location`
- `type`
- `quantile`
- `value`

No additional columns are allowed.

Each row in the file is either a point or quantile forecast for a location on a 
particular date for a particular target. 


### `forecast_date`

Values in the `forecast_date` column must be a date in the format

    YYYY-MM-DD
    
This is the date on which the submitted forecast were available.
This will typically be the date on which the computation finishes running and 
produces the standard formatted file. 
`forecast_date` should correspond and be redundant with the date in the filename, 
but is included here by request from some analysts. 
We will enforce that the `forecast_date` for a file must be either the date on 
which the file was submitted to the repository or the previous day. 
Exceptions will be made for legitimate extenuating circumstances.


### `target`

Values in the `target` column must be a character (string) and be one of the 
following specific targets:

- "N wk ahead cum death"  where N is a number between 1 and  20
- "N wk ahead inc death"  where N is a number between 1 and  20
- "N wk ahead inc case"  where N is a number between 1 and  8
- "N day ahead inc hosp"  where N is a number between 0 and 130

For county locations, the only target should be "N wk ahead inc case".

For week-ahead forecasts, we will use the specification of epidemiological weeks (EWs) [defined by the US CDC](https://wwwn.cdc.gov/nndss/document/MMWR_Week_overview.pdf) which run
Sunday through Saturday.
There are standard software packages to convert from dates to epidemic weeks and vice versa. E.g. [MMWRweek](https://cran.r-project.org/web/packages/MMWRweek/) for R and [pymmwr](https://pypi.org/project/pymmwr/) and [epiweeks](https://pypi.org/project/epiweeks/) for python.

We have created [a csv file](../template/covid19-death-forecast-dates.csv) describing forecast collection dates and dates for which forecasts refer to can be found.


For week-ahead forecasts with `forecast_date` of Sunday or Monday of EW12, 
a 1 week ahead forecast corresponds to EW12 and should have `target_end_date` of 
the Saturday of EW12. For week-ahead forecasts with `forecast_date` of Tuesday 
through Saturday of EW12, a 1 week ahead forecast corresponds to EW13 and should 
have `target_end_date` of the Saturday of EW13. 

#### N wk ahead cum death

This target is the cumulative number of deaths predicted by the model up to 
and including N weeks after `forecast_date`. 

A week-ahead forecast should represent the cumulative number of deaths reported 
on the Saturday of a given epiweek.

Predictions for this target will be evaluated compared to the cumulative of the
number of new reported cases, as recorded by 
[JHU CSSE](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv).


#### N wk ahead inc death

This target is the incident (weekly) number of deaths predicted by the model 
during the week that is N weeks after `forecast_date`. 

A week-ahead forecast should represent the total number of new deaths 
reported during a given epiweek (from Sunday through Saturday, inclusive).

Predictions for this target will be evaluated compared to the number of new 
reported cases, as recorded by 
[JHU CSSE](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv).


#### N wk ahead inc case

This target is the incident (weekly) number of cases predicted by the model 
during the week that is N weeks after `forecast_date`. 

A week-ahead forecast should represent the total number of new cases 
reported during a given epiweek (from Sunday through Saturday, inclusive).

Predictions for this target will be evaluated compared to the number of new 
reported cases, as recorded by 
[JHU CSSE](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv).


#### N day ahead inc hosp

This target is the number of new daily hospitalizations predicted by the
model on day N after `forecast_date`.

As an example, for day-ahead forecasts with a `forecast_date` of a Monday, 
a 1 day ahead inc hosp forecast corresponds to the number of incident 
hospitalizations on Tuesday, 2 day ahead to Wednesday, etc.... 

Currently there is no "gold standard" for hospitalization data. 



#### REMOVED targets

On 2020-06-06, these targets were removed:

- N day ahead inc death
- N day ahead cum death




### `target_end_date`

Values in the `target_end_date` column must be a date in the format

    YYYY-MM-DD
    
This is the date for the forecast `target`. 
For "# day" targets, `target_end_date` will be # days after `forecast_date`.
For "# wk" targets, `target_end_date` will be the Saturday at the end of the 
week time period.


### `location`

Values in the `location` column must be one of the "locations" in this
[FIPS numeric code file](../data-locations/locations.csv) which includes 
numeric FIPS codes for U.S. states, counties, territories, and districts as
well as "US" for national forecasts. 

Please note that when writing FIPS codes, they should be written in as a
character string to preserve any leading zeroes.


### `type`

Values in the `type` column are either

- "point" or
- "quantile". 

This value indicates whether that row corresponds to a point forecast or a 
quantile forecast. 
Point forecasts are used in visualization while quantile forecasts are used in
visualization and in ensemble construction.

**Forecasts must include exactly 1 "point" forecast for every location-target
pair.**


### `quantile`

Values in the `quantile` column are either "NA" (if `type` is "point") or 
a quantile in the format

    0.###
    
For quantile forecasts, this value indicates the quantile for the `value`
in this row. 

Teams should provide the following 23 quantiles:

```{r, echo = TRUE}
c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)
```

for all `target`s except "N wk ahead inc case" target.
For the "N wk ahead inc case" target, teams should provide the following 6 
quantiles: 

```{r, echo = TRUE}
c(0.025, 0.100, 0.250, 0.500, 0.750, 0.900, 0.975)
```


### `value`

Values in the `value` column are non-negative numbers indicating the "point" or 
"quantile" prediction for this row. 
For a "point" prediction, `value` is simply the value of that point prediction
for the `target` and `location` associated with that row. 
For a "quantile" prediction, `value` is the inverse of the cumulative 
distribution function (CDF)
for the `target`, `location`, and `quantile` associated with that row.

An example inverse CDF is below.

```{r example_inverse_cdf, echo=FALSE, fig.path="./"}
mu = log(50000)
q = 0.800
value = qlnorm(q, mu)
curve(qlnorm(x, mu), 0.1, 0.9,
      xlab = "quantile", 
      ylab = "inverse CDF",
      main = paste("Example `value` at quantile", q))
segments(q, 0, q, value)
segments(0, value, q, value)
axis(2, at = value, "value", las = 2)
```



## Forecast validation

To ensure proper data formatting, 
pull requests for new data in data-processed/ will be automatically run.



### Pull request forecast validation

When a pull request is submitted, 
the data are validated through [Travis CI](https://travis-ci.org/) which runs
the tests in [test_formatting.py](../code/validation/test_formatting.py).
The intent for these tests are to validate the requirements above and 
specifically enumerated [on the wiki](https://github.com/reichlab/covid19-forecast-hub/wiki/Validation-Checks#current-validation-checks).
Please [let us know](https://github.com/reichlab/covid19-forecast-hub/issues) 
if the wiki is inaccurate.

If the pull request fails, please 
[follow these instructions](https://github.com/reichlab/covid19-forecast-hub/wiki/Troubleshooting-Pull-Requests)
for details on how to troubleshoot.

#### Run checks locally

To run these checks locally rather than waiting for the results from a pull request, follow
[these instructions](https://github.com/reichlab/covid19-forecast-hub/wiki/Validation-Checks#running-validations-locally).


### R validation checks

If you cannot get the python checks to run, 
you can use [these instructions](R_forecast_file_validation.md) to run some
checks in R. 
These checks are no longer maintained, but may still be of use to teams working with R.


## Data visualization

If you want to visualize your forecasts, 
you can use our [R shiny app](./explore_processed_data.R) 
to visualize your forecast by running

```{r, eval=FALSE}
source("explore_processed_data.R")
shinyApp(ui = ui, server = server)
```

from within the [data-processed/](./) folder.
This is mainly an internal tool we use to help us know what forecasts are in 
the repository. 
Thus, it is provided as-is within no warranty. 



