---
title: "aa_state_weekly_report_county_code"
output: html_document
---

```{r accuracy of models past 6 weeks-county}

#inc cases for past 6 weeks
inc_case_p6_c <- data.frame(inc_case_all_points %>%
      rename(location=fips,point=value)%>%
      select(target_end_date,location,location_name,point) %>%
    filter(location %in% county,target_end_date %in% last_6_saturdays))
inc_case_p6_u_c <- unique(inc_case_p6_c)

# submit query for forecasts past 6 weeks-using covidhubutils
inc_case_forecast_p6_c<-load_forecasts(models=c("COVIDhub-ensemble"),dates =last_6_mondays,locations =county,  types = c("quantile"),targets = c(inc_case_targets),  verbose = FALSE)

## join inc cases with forecast
inc_case_all_p6_c  <- inc_case_forecast_p6_c %>%
  dplyr::full_join(inc_case_p6_u_c,inc_case_forecast_p6_c,by=c("target_end_date", "location"))%>%
  dplyr::filter(quantile %in% quantiles_to_plot,target_end_date<=this_monday) %>%
  dplyr::mutate(endpoint_type = ifelse(quantile < 0.5, 'lower', 'upper'),
    alpha = ifelse(endpoint_type == 'lower',
      format(2*quantile, digits=3, nsmall=3),
      format(2*(1-quantile), digits=3, nsmall=3)),
    `Prediction Interval` = fct_rev(paste0((1-as.numeric(alpha))*100, "%"))
  ) %>%
  dplyr::filter(alpha != "1.000") %>%
  dplyr::select(-quantile) %>%
  tidyr::pivot_wider(names_from='endpoint_type', values_from='value')

#identify if inc_case with PI
inc_case_all_p6x_c<-inc_case_all_p6_c %>%
  select(target_end_date,location,alpha,lower,upper,point )%>%
dplyr::mutate(inc_case_inPI = ifelse(point>=lower & point<=upper, 1, 0)) %>%
dplyr::mutate(inc_case_inPI_low = ifelse(point<=lower, 1, 0)) %>%
dplyr::mutate(inc_case_inPI_high = ifelse(point>=upper, 1, 0))

# proportion PI covered truth
pcounty_inc_case_PI_c<-inc_case_all_p6x_c %>% group_by(alpha) %>%
  filter(location %in% county) %>%
  summarise_at(vars(inc_case_inPI,inc_case_inPI_low,inc_case_inPI_high), list(PI_cover=mean,PI_coverl=mean,PI_coverh=mean), na.rm=TRUE) %>% 
    dplyr::mutate(PI_coverp = as.numeric(substr(inc_case_inPI_PI_cover*100, 0,2))) %>% 
    dplyr::mutate(PI_coverpl = as.numeric(substr(inc_case_inPI_low_PI_coverl*100, 0,2))) %>%   
    dplyr::mutate(PI_coverph = as.numeric(substr(inc_case_inPI_high_PI_coverh*100, 0,2)))  %>%
    dplyr::mutate(bias = ifelse(PI_coverpl<=PI_coverph, 'lower', 'higher'))

C_PI95_cover_c<-pcounty_inc_case_PI_c %>%
  filter(alpha=="0.050")
C_PI95_covern_c<-paste0(C_PI95_cover_c$PI_coverp)
PI95_bias_c<-paste0(C_PI95_cover_c$bias)

C_PI50_cover_c<-pcounty_inc_case_PI_c %>%
  filter(alpha=="0.500") %>% 
dplyr::mutate(pi_in = ifelse(PI_coverp>=max(PI_coverph,PI_coverpl), 'yes', 'no'))
C_PI50_covern_c<-paste0(C_PI50_cover_c$PI_coverp)
C_PI50_bias_c<-paste0(C_PI50_cover_c$bias)

bias_note_c_c<-if (C_PI50_cover_c$pi_in == 'no') {
       paste0("The forecasts on average over the last 6 weeks predicted values that were ", C_PI50_bias_c, " than the observed data.")
       } else  {
       paste0("")
       }



```

## County level


County-level forecasts are only made for cases. County-level case forecasts are presented below.   

Cases reported in unknown counties or by people who reside outside of the state are not included in Table 3,  but are included in the state-level data.   

`r exclude`


```{r R6-class}
# make R6 class that represents a shared data frame
inc_case_all_points_df_full <-  SharedData$new(inc_case_all_points %>%
    dplyr:: filter(fips != 'US' & fips != state) ,
  ~fips,
  group = "county")

inc_case_all_points_df_prediction<-SharedData$new(
  inc_case_all_points %>%
    dplyr::filter((fips != 'US' & fips != state) & !(model=="COVIDhub-ensemble" & target_end_date <= this_monday)),
  ~fips,
  group = "county")

inc_case_quantiles_df <- SharedData$new(
  inc_case_quantiles %>%
    dplyr::filter(fips != 'US' & fips != state),
  ~fips,
  group = "county")


# make ggplot object
county_plot <- ggplot() +

  # dummy points just for tooltip
  geom_point(data = inc_case_quantiles_df,
             mapping = aes(x = target_end_date,
      y=lower, fill = `Prediction Interval`,
      # make the dots transparent
      stroke = 0, shape = '.', alpha=I(0.1), text = text ),
      show.legend = FALSE) +

  # dummy points just for tooltip
  geom_point(data = inc_case_quantiles_df,
             mapping = aes(x = target_end_date,
      y=upper, fill = `Prediction Interval`,
      # make the dots transparent
      stroke = 0,shape = '.', alpha=I(0.1),text = text ),
      show.legend = FALSE)+
  # color for dummy points
  scale_fill_manual(
                    values = blues[1:(length(blues)-1)])+

  geom_ribbon(data = inc_case_quantiles_df,
    mapping = aes(x = target_end_date,
      ymin=lower, ymax=upper,
      fill=`Prediction Interval`)) +
  geom_line(data=inc_case_all_points_df_full,
    mapping = aes(x = target_end_date, y = value, color = model)) +
  geom_point(data=inc_case_all_points_df_prediction,
    mapping = aes(x = target_end_date, y = value, color = model,
                  text = paste(sprintf("%s", location_name),
                               sprintf("For week ending: %s", target_end_date),
                               ifelse(model == "COVIDhub-ensemble",
                                      sprintf("Incident cases (forecasted): %s",value),
                                      sprintf("Incident cases (observed): %s",value)),
                               sep = '<br />')
                  )) +
  scale_fill_manual(values = blues[1:(length(blues)-1)]) +
  scale_color_manual(values = c("black", tail(blues,1))) +
  scale_x_date(name = NULL, date_breaks="4 month",date_labels = "%b %d %Y", date_minor_breaks = "1 month") +
  ylab("incident cases") +
  theme()
```

```{r plot-county-cases}
# create caption
plot3_cap <- paste0("Weekly reported COVID-19 cases in ", state_name," counties: observed and forecasted")

# convert ggplot object to plotly object and add titles and captions
county_plot <- plotly::ggplotly(county_plot, dynamicTicks = "y",tooltip = "text") %>%
    # add title
    layout(annotations=list(text=plot3_cap,
                            xref="paper",
                            # location on x axis
                            x=0.5, xshift = -40,
                            yref="paper",
                            # location on y axis
                            y=1, yshift=30,
                            showarrow=FALSE,
                            # font size
                            font=list(size=12)),
           # turn on legends
           showlegend = TRUE,
           # put legend below the graph horizontally
           legend = list(orientation = "h", x = 0.5, y = -0.2)) %>%
  # add caption
  layout(annotations=list(text="source: JHU CSSE (observed data), COVID-19 Forecast Hub (forecasts)",
                          xref="paper",
                          # location on x axis
                          x=1, xshift = 0,
                          yref="paper",
                          # location on y axis
                          y=0, yshift=-40,
                          showarrow=FALSE,
                          # font size
                          font=list(size=10)),
         # remove date axis label
         xaxis = list(title = ''))

# change variable name in tooltip
for (i in 1:length(county_plot$x$data)){
  # change legend names
  if (!is.null(county_plot$x$data[[i]]$name)){

    county_plot$x$data[[i]]$name =
      gsub("\\(","",str_split(county_plot$x$data[[i]]$name,",")[[1]][1])

    county_plot$x$data[[i]]$text =
      str_replace(county_plot$x$data[[i]]$text, ".crossTalkKey", "fips")
  }

  # turn off lengend for point layer
  if (i <= 6){
      county_plot$x$data[[i]]$showlegend<- FALSE
  }
  # change ribbon layer legend text
  else if ( i %in% c(7,8,9)){
      county_plot$x$data[[i]]$name <- paste(county_plot$x$data[[i]]$name, "Prediction Interval")
  }

  # change line layer legend text and symbol
  else if (i == 10| i == 11){
      county_plot$x$data[[i]]$name <- ifelse (grepl("observed", county_plot$x$data[[i]]$name , fixed = TRUE), "Observed Data (JHU)", county_plot$x$data[[i]]$name)
      county_plot$x$data[[i]]$mode <- 'lines+markers'
    }

}

# remove legend text from ggplot
county_plot$x$layout$annotations[[1]]$text <- "  "
# change legend font size
county_plot$x$layout$legend$font$size<-8

# show filter and graph
bscols(widths = c(3,NA),

       # filter
       filter_select(id= "location_name", label = "Select a County", sharedData = inc_case_all_points_df_full, ~location_name, multiple = FALSE),

       # graph object
       county_plot
)

```


The sortable and searchable table below shows the total number of reported COVID-19 cases at the county level for `r state_name` as of Saturday,`r format(last_5_saturdays[5], "%B %d, %Y")`  ("Total COVID-19 Cases") as well as the rate of reported COVID-19 cases in the population (standardized per 100,000 population) over the last two weeks and over the next two weeks.

For county-level case forecasts made over the last 6 weeks in `r state_name` for which observed data are now available, the 95% prediction intervals covered the truth `r C_PI95_covern`% of the time and the 50% prediction intervals covered the truth `r C_PI50_covern`% of the time.  `r bias_note_c_c` 

Looking at the rates allows for easier comparison across counties, where you can see which counties have had or are predicted to have propoportionally higher rates in comparison to other counties. These tables calculate the weekly case count and rate of reported COVID-19 cases in the population (standardized per 100,000 population) over the last week(`r format(last_5_saturdays[4]+1, "%B %d, %Y")` to `r format(last_5_saturdays[5], "%B %d, %Y")`) and two weeks ahead (`r format(last_5_saturdays[5]+1, "%B %d, %Y")` to `r format(last_5_saturdays[5]+7, "%B %d, %Y")`).

When interpreting probability of an increase, it's important to note that the increase or decrease is relative to the location-specific recent observed incidence, which varies across locations.


```{r process-ensemble-data-cases for counties}

ensemble_pointdat1  <- dat  %>%
  filter(grepl('inc case', target)) %>%
  filter(class=="point") %>%
  filter(fips %in% county) %>%
  select(fips, target, value, timezero)

wide_point_dat1 <- spread(ensemble_pointdat1, target, value) %>%
  left_join(recent_inc_case_totals) %>%
  mutate(next_1wk_cases = `1 wk ahead inc case`,
    pop_x_1k = round(Population/1000),
    last_1wk_cases_rate = round(last_1wk_cases/Population*100000/7,3),
    last_2wk_cases_rate = round(last_2wk_cases/Population*100000/14,3),
    next_1wk_cases_rate = round(next_1wk_cases/Population*100000/7, 3),
    diff_wk_cases_rate = round(next_1wk_cases_rate - last_1wk_cases_rate, 3)) %>%
  select(location_name.x, Population,
    last_1wk_cases,last_2wk_cases, next_1wk_cases,  last_1wk_cases_rate,  last_2wk_cases_rate,next_1wk_cases_rate, diff_wk_cases_rate)



#filter find first quantile where predicting future 2 weeks >= previous week count
ensemble_quantdat1 <- dat %>%
  filter(target == "1 wk ahead inc case") %>%
  filter(class == "quantile")  

wide_quant_dat1 <- spread(ensemble_quantdat1, target, value) %>%
  left_join(recent_inc_case_totals %>% select(fips, location_name.x, last_1wk_cases)) %>%
  mutate(next_1wk_cases = `1 wk ahead inc case`)

quant_cutoff1 <- wide_quant_dat1 %>%
  filter(next_1wk_cases >= last_1wk_cases) %>%
  mutate(prob_case_up = 1-quantile) %>%
select( quantile, location_name.x, prob_case_up,class)

quant_cutoff1min<-quant_cutoff1[!duplicated(quant_cutoff1$location_name.x),]

wide_point_dat1p <- merge(wide_point_dat1, quant_cutoff1min, by="location_name.x")


```


  
```{r make-datatable-inc-case-rates county}

## color for rates
case_rate_max_1wk <- max(c(wide_point_dat1$last_1wk_cases_rate))
brks <- seq(0, case_rate_max_1wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

## colors for rate difference
# case_rate_diff_2wk <- max(abs(wide_point_dat1$diff_2wk_cases_rate))
# brks1 <- seq(-case_rate_diff_2wk, case_rate_diff_2wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
# clrs1 <- colorRampPalette(colors = rev(RColorBrewer::brewer.pal(n=3, "RdBu")))(length(brks1)+1)

table2_cap <- paste0("Table 2:",  state_name, " county level observed and predicted COVID-19 cases for the previous week  and one week ahead.")

table2_Observed <- paste0("Observed: (", format(last_5_saturdays[4]+1, "%m-%d-%Y") ," to ",format(last_5_saturdays[5], "%m-%d-%Y"),")")

table2_Predicted <- paste0("Predicted: (", format(last_5_saturdays[5]+1, "%m-%d-%Y") ," to ",format(last_5_saturdays[5]+7, "%m-%d-%Y"),")")


# a custom table container
sketch_tab3 = htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(rowspan = 2, "County"),
      th(rowspan = 2, "Population"),
      th(colspan = 2, table2_Observed),
      th(colspan = 2, table2_Predicted),
      th(rowspan=2,"Case rate difference")
    ),
    tr(
 lapply((c("Weekly count (Obs)","Rate per 100K (Obs)","Weekly count (Pred)","Rate per 100K (Pred)")), th)))))

datatable(wide_point_dat1p %>%
    select(location_name.x, Population,
       last_1wk_cases,
      last_1wk_cases_rate, next_1wk_cases, next_1wk_cases_rate, diff_wk_cases_rate) %>%
    arrange(desc(last_1wk_cases_rate)),
  caption = table2_cap,
  options = list(
    autoWidth = TRUE,
    columnDefs = list(list(width = '100px', targets = c(0, 1, 2, 3, 4, 5,6))),
  pageLength = 25),

  #width=paste0(c(10, 100, 100, 100), 'px'),
  rownames=FALSE,
  colnames=c('county'='location_name.x',
    # 'Population, \'000'='pop_x_1k',
    'Weekly Case count (Obs)'='last_1wk_cases',
    'Rate per 100K (Obs)'='last_1wk_cases_rate',
    'Weekly Case count (Pred)'='next_1wk_cases',
    'Rate per 100K (Pred)'='next_1wk_cases_rate',
     'Death rate difference'='diff_wk_cases_rate'),
     container=sketch_tab3) %>%
  ## formatStyle("Rate per 100K", backgroundColor = styleInterval(brks, clrs)) %>%
  ## formatStyle("Rate per 100K", backgroundColor = styleInterval(brks, clrs))  %>%
  # formatStyle('Case rate difference', backgroundColor = styleInterval(brks1, clrs1))  %>%
  formatCurrency('Population',currency = "", interval = 3, mark = ",", digits=0)%>%
  formatCurrency('Weekly Case count (Obs)',currency = "", interval = 3, mark = ",", digits=0)%>%
  formatCurrency('Rate per 100K (Pred)',currency = "", interval = 3, mark = ",", digits=3)%>%
  formatCurrency('Rate per 100K (Obs)',currency = "", interval = 3, mark = ",", digits=3)%>%
 formatCurrency('Weekly Case count (Pred)',currency = "", interval = 3, mark = ",", digits=0)
```