---
title: "COVID-19 US Weekly Forecast Summary"
author: "[The COVID-19 Forecast Hub Team](https://covid19forecasthub.org/doc/)<br><br>https://covid19forecasthub.org/"
date: "report generated `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
params:
  conn: NA
  quiet: false
---
<!-- code to run rmarkdown::render(input="./vignettes/covidHubUtils-overview.Rmd") -->

<!-- Code for adding logo at the top -->

<!-- <script> -->
<!--   $(document).ready(function() { -->
<!--     $('#TOC').parent().prepend('<div id=\"nav_logo\"><a href=\"https://covid19forecasthub.org/\" target=\"_blank\"><img src=\"https://github.com/reichlab/covid19-forecast-hub-web/raw/master/images/forecast-hub-logo_DARKBLUE.png\"></a></div>'); -->
<!--   }); -->
<!-- </script> -->

<!-- <style> -->
<!-- #nav_logo { -->
<!--   width: 100%; -->
<!--   margin-top: 20px; -->
<!-- } -->

<!-- #TOC { -->
<!--   background: url("https://github.com/reichlab/covid19-forecast-hub-web/raw/master/images/forecast-hub-logo_DARKBLUE-20px-padding.png"); -->
<!--   background-size: contain; -->
<!--   padding-top: 80px !important; -->
<!--   background-repeat: no-repeat; -->
<!-- } -->
<!-- </style> -->
<!-- </style> -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(lubridate)
library(DT)
library(zoltr) ## devtools::install_github("reichlab/zoltr")
library(scico)
library(tidyverse)
library(htmltools)
library(covidHubUtils)
library (zoo)
theme_set(theme_bw())
```

```{r zoltar-setup}
zoltar_connection <- NA

# make a new connection if no connection given
if (!inherits(params$conn, "ZoltarConnection")) {
  ## connect to Zoltar
  zoltar_connection <<- new_connection()
  
  # try to connect to Zoltar 5 times; if all fail, fail the script
  num_tries <- 0
  success <- FALSE
  while(num_tries < 5 && !success) {
    tryCatch(
      # try to authenticate
      {
        zoltar_authenticate(
          zoltar_connection,
          Sys.getenv("Z_USERNAME"),
          Sys.getenv("Z_PASSWORD")
        )
        # <<- superassignment: should only do if preceding scope have
        # such a variable!
        # this statement is reached only if authentication is successful
        success <<- TRUE
      },
      # authentication failed! retry
      error = function(c) {
        message(sprintf("Zoltar connection failed! %d retries remaining...", num_tries))
      },
      # add one to number of retries
      finally = function(c) {
        # <<- superassignment: should only do if preceding scope have
        # such a variable!
        num_tries <<- num_tries + 1
      }
    )
  }
  if (!success) {
    stop("Could not make connection to Zoltar after 5 tries")
  }
# use connection if given
} else {
  zoltar_connection <<- params$conn
}


## construct Zoltar query
project_url <- "https://www.zoltardata.com/api/project/44/"
```


```{r get-date-boundaries}
 # use fixed date
# next_saturday <-  as.Date("2021-06-12")
next_saturday <- as.Date(calc_target_week_end_date(today(), horizon = 0))

saturday_4_wk_ahead <- next_saturday + 7*3
saturday_4_wk_ahead_txt <- format(saturday_4_wk_ahead, "%B %d, %Y")
saturday_2_wk_ahead <- next_saturday + 7*1
saturday_2_wk_ahead_txt <- format(saturday_2_wk_ahead, "%B %d, %Y")
saturday_1_wk_ahead_txt <- format(next_saturday, "%B %d, %Y")
last_5_saturdays <- next_saturday - 7*c(5:1)
this_monday <- next_saturday - 5
next_2saturday <- next_saturday +7
wk4_monday <- this_monday + 28
wk2_monday <- this_monday + 14
```


# Background
This report provides a brief summary of the weekly ensemble forecast from the [COVID-19 Forecast Hub](https://covid19forecasthub.org/) based on forecasts submitted on `r format(this_monday, "%B %d, %Y")`. In collaboration with the US CDC, our team aggregates COVID-19 forecasts from dozens of teams around the globe. Typically on Wednesday of each week, a summary of the week's forecasts from the COVID-19 Forecast Hub appear on the [official CDC COVID-19 forecasting page](https://www.cdc.gov/coronavirus/2019-ncov/covid-data/forecasting-us.html).


```{r nmodels-this-week}
possible_timezeroes <- seq.Date(this_monday, this_monday-6, by="-1 day")
this_week_timezeroes <- timezeros(zoltar_connection, project_url) %>%
  filter(timezero_date %in% possible_timezeroes) %>%
  pull(timezero_date) %>% sort.default()

models_this_week<-load_forecasts(
        # models=c(),
        dates = this_week_timezeroes,
        types = c("point"),
        targets = c("1 wk ahead cum death", "1 wk ahead inc death", "1 wk ahead inc case", "1 day ahead inc hosp"),
        verbose = FALSE)%>%
  pull(model) %>% sort.default()%>% 
  unique()
nmodels_this_week <- length(models_this_week)


```


Every week, teams submit their forecasts to the COVID-19 Forecast Hub. 
This past week, `r nmodels_this_week` models were submitted. 

Each Monday evening or Tuesday morning, we combine the most recent forecasts from each team into a single "ensemble" forecast of reported COVID-19 cases at the county, state, and national level and deaths at the state and national level. At the moment, we only generate ensemble forecasts for up to four weeks into the future, as [the available evidence](https://www.medrxiv.org/content/10.1101/2021.02.03.21250974v1) suggests that models are less accurate at longer forecast horizons.

An archive of weekly reports from the COVID-19 Forecast Hub can be found at [this page](https://covid19forecasthub.org/doc/reports/). 


```{r count-models}
## how many models in inc_death ensemble?
inc_death_models <- read_csv(paste0("../../ensemble-metadata/", this_monday, "-inc_death-model-weights.csv")) %>%
 select(-locations) %>%
  apply(MARGIN = 2, FUN=function(x) sum(x))
n_inc_death_models <- sum(inc_death_models>0)

## how many models in cum_death ensemble?
cum_death_models <- read_csv(paste0("../../ensemble-metadata/", this_monday, "-cum_death-model-weights.csv")) %>%
  select(-locations) %>%
  apply(MARGIN = 2, FUN=function(x) sum(x))
n_cum_death_models <- sum(cum_death_models>0)

## how many models in inc_case ensemble?
inc_case_models <- read_csv(paste0("../../ensemble-metadata/", this_monday, "-inc_case-model-weights.csv"))%>%
  select(-locations) %>%
  apply(MARGIN = 2, FUN=function(x) sum(x))
n_inc_case_models <- sum(inc_case_models>0)

## how many models in inc_hosp ensemble?
inc_hosp_models <- read_csv(paste0("../../ensemble-metadata/", this_monday, "-inc_hosp-model-weights.csv"))%>%
  select(-locations) %>%
  apply(MARGIN = 2, FUN=function(x) sum(x))
n_inc_hosp_models <- sum(inc_hosp_models>0)

n_unique_models <- length(unique(c(names(inc_death_models)[inc_death_models>0],
  names(cum_death_models)[cum_death_models>0],
  names(inc_case_models)[inc_case_models>0],
  names(inc_hosp_models)[inc_hosp_models>0])))
  
```

```{r aux-data}
locs <- hub_locations %>%
  rename(Population = population)
all_states <-locs$fips[2:58]
```


```{r download-ensemble-data}
inc_death_targets <- paste(1:4, "wk ahead inc death")
cum_death_targets <- paste(1:4, "wk ahead cum death")
inc_case_targets <- paste(1:1, "wk ahead inc case")
inc_hosp_targets <- paste(1:14, "day ahead inc hosp")

# submit query with covidHubUtils

##deaths,cases
dat<-load_forecasts(
        models=c("COVIDhub-ensemble"),
        dates = this_monday, 
        types = c("point", "quantile"),
        targets = c(inc_death_targets, cum_death_targets,inc_case_targets),
       location=c("US",all_states),
       verbose = FALSE) %>%
     rename(fips=location, timezero=forecast_date, class=type, week_ahead=horizon)%>%
     mutate(target=paste(week_ahead,target_variable,sep=" wk ahead "))%>%
     select(model, timezero, fips, target, class, quantile, value) %>%
     # create rate variable and week-ahead
     mutate(week_ahead = as.numeric(substr(target, 0,1)),
     ## recreates the target_end_date from GitHub
     target_end_date = as.Date(calc_target_week_end_date(timezero, horizon = week_ahead)))


##hospitailization

dat_hosp<-load_forecasts(
        models=c("COVIDhub-ensemble"),
        dates = this_monday, 
        types = c("point", "quantile"),
        targets = paste(inc_hosp_targets),
       location=c("US",all_states),
       verbose = FALSE,
       source = "zoltar") %>%
     rename(fips=location, timezero=forecast_date, class=type, day_ahead=horizon)%>%
     mutate(target=paste(day_ahead,target_variable,sep=" day ahead "))%>%
     select(model, timezero, fips, target, class, quantile, value, day_ahead) %>%
     # create rate variable and week-ahead
     mutate(day_ahead = as.numeric(substr(target, 0,2)),
     ## recreates the target_end_date from GitHub
     target_end_date = as.Date(timezero + day_ahead))    
```


# COVID-19 Forecasts 

## National level {.tabset .tabset-fade}


This week, our ensemble combined forecasts from `r n_unique_models` different models. 

```{r us-summary}
##deaths
us_cum_deaths <- dat %>% 
  filter(fips=="US", target=="4 wk ahead cum death", class=="point") %>% 
  pull(value) %>% 
  round(-2) %>% 
  format(big.mark = ",")
us_cum_deaths_wk_pi<- dat %>% 
  filter(fips=="US", target=="4 wk ahead cum death", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% 
  format(big.mark = ",")
us_inc_death_range <- dat %>% 
  filter(fips=="US", target %in% inc_death_targets, class=="point") %>% 
  pull(value) %>% range() %>%
  round(-2) %>% format(big.mark = ",")

us_inc_death_wk_pi_round <- dat %>% 
  filter(fips=="US", target == "4 wk ahead inc death", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% sort() %>%
  round(-2) %>% format(big.mark = ",")

us_inc_death_wk_pi <- dat %>% 
  filter(fips=="US", target == "4 wk ahead inc death", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% sort() %>% 
  format(big.mark = ",")

##cases
us_inc_case_range <- dat %>% 
  filter(fips=="US", target %in% inc_case_targets, class=="point") %>% 
  pull(value) %>% range() %>%
  round(-2) %>% format(big.mark = ",")

us_inc_case_wk_pi_round <- dat %>% 
  filter(fips=="US", target == "1 wk ahead inc case", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% sort() %>%
  round(-2) %>% format(big.mark = ",")

us_inc_case_wk_pi <- dat %>% 
  filter(fips=="US", target == "1 wk ahead inc case", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% sort() %>% 
  format(big.mark = ",")

##hospitalization
us_inc_hosp_range <- dat_hosp %>% 
  filter(fips=="US", target %in% inc_hosp_targets, class=="point") %>% 
  pull(value) %>% range() %>%
  round(-2) %>% format(big.mark = ",")

us_inc_hosp_wk_pi_round <- dat_hosp %>% 
  filter(fips=="US", target == "14 day ahead inc hosp", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% sort() %>%
  round(-2) %>% format(big.mark = ",")

us_inc_hosp_wk_pi <- dat_hosp %>% 
  filter(fips=="US", target == "14 day ahead inc hosp", quantile %in% c(0.025, 0.975)) %>% 
  pull(value) %>% sort() %>% 
  format(big.mark = ",")

```

At the national level, the ensemble model predicts that weekly totals of observed deaths in each of the next four weeks will be between `r us_inc_death_range[1]` and `r us_inc_death_range[2]` deaths with around `r us_cum_deaths` deaths by `r saturday_4_wk_ahead_txt` (95% prediction interval: `r us_cum_deaths_wk_pi[1]` - `r us_cum_deaths_wk_pi[2]`). 

For the week ending `r saturday_4_wk_ahead_txt`, the ensemble forecasts that reported COVID-19 deaths in the US will be between `r us_inc_death_wk_pi_round[1]` and `r us_inc_death_wk_pi_round[2]` (95% prediction interval: `r us_inc_death_wk_pi[1]` - `r us_inc_death_wk_pi[2]`).  

As of September 28, 2021 the ensemble forecast only reports one-week ahead forecasts for cases and 14 day ahead forecasts for hospitalizations, due to persistent large inaccuracies observed when forecasting beyond that. 

For `r format(wk2_monday, "%B %d, %Y")` COVID-19 daily hospitalizations in the US will be between `r us_inc_hosp_wk_pi_round[1]` and `r us_inc_hosp_wk_pi_round[2]` (95% prediction interval: `r us_inc_hosp_wk_pi[1]` - `r us_inc_hosp_wk_pi[2]`).

For the week ending `r saturday_1_wk_ahead_txt`, the ensemble forecasts that reported  COVID-19 cases in the US will be between `r us_inc_case_wk_pi_round[1]` and `r us_inc_case_wk_pi_round[2]` (95% prediction interval: `r us_inc_case_wk_pi[1]` - `r us_inc_case_wk_pi[2]`). 



<!-- Throughought most of July, models have in general shown broad agreement about the trajectory of the outbreak over the coming weeks. However, the recent surge in cases has left models with quite different interpretations about what the next few weeks hold in terms of how many reported deaths from COVID-19 we will see.  -->
You can explore the full set of models, including their forecasts for past weeks online at the [Forecast Hub interactive visualization](https://viz.covid19forecasthub.org/).

### Deaths

```{r make-US-inc-death-plot}
quantiles_to_plot <- c(0.025, 0.1, 0.25, 0.75, 0.9, 0.975)

blues <- RColorBrewer::brewer.pal(n=length(quantiles_to_plot)/2+1, "Blues")

inc_death_forecast <- dat %>%
  filter(target %in% inc_death_targets)

# get full inc death truth for plotting
inc_death_truth <- load_truth(
                       truth_source = "JHU",
                       target_variable = "inc death")%>%
                    rename(fips = location) %>%
mutate(model = "observed data (JHU)") %>%     #maybe also deleting this line
                    left_join(locs, by=c("fips")) %>%
                    filter(fips %in% unique(inc_death_forecast$fips), target_end_date <= last_5_saturdays[5])



inc_death_all_points <- inc_death_truth %>%
  bind_rows(filter(inc_death_forecast, class=="point")) %>%
  bind_rows(filter(inc_death_truth, target_end_date==last_5_saturdays[5]) %>% mutate(model="COVIDhub-ensemble")) %>%
  mutate(model = relevel(factor(model), ref="observed data (JHU)"))

# inc death data for code of uncertainty
dummy_inc_death <- tibble(
  quantile = quantiles_to_plot, 
  target_end_date=last_5_saturdays[5]) %>%
  right_join(inc_death_all_points %>%
      select(-quantile) %>%
      filter(target_end_date == last_5_saturdays[5]))

inc_death_quantiles <- inc_death_forecast %>%
  dplyr::filter(class=="quantile") %>%
  bind_rows(dummy_inc_death) %>%
  dplyr::filter(quantile %in% quantiles_to_plot) %>%
  dplyr::mutate(endpoint_type = ifelse(quantile < 0.5, 'lower', 'upper'),
    alpha = ifelse(endpoint_type == 'lower',
      format(2*quantile, digits=3, nsmall=3),
      format(2*(1-quantile), digits=3, nsmall=3)),
    `Prediction Interval` = fct_rev(paste0((1-as.numeric(alpha))*100, "%"))
  ) %>%
  dplyr::filter(alpha != "1.000") %>%
  dplyr::select(-quantile) %>%
  tidyr::pivot_wider(names_from='endpoint_type', values_from='value')

ggplot() +
  geom_ribbon(data = inc_death_quantiles %>% dplyr::filter(fips=="US"), 
      mapping = aes(x = target_end_date,
      ymin=as.numeric(lower), ymax=as.numeric(upper),
      fill=`Prediction Interval`)) +
  geom_line(data=inc_death_all_points %>%
      dplyr::filter(fips == "US"),
    mapping = aes(x = target_end_date, y = value, color = model)) +
  geom_point(data=inc_death_all_points %>%
      dplyr::filter(fips == "US", !(model=="COVIDhub-ensemble" & target_end_date <= this_monday)),
    mapping = aes(x = target_end_date, y = value, color = model)) +
  scale_fill_manual(values = blues[1:(length(blues)-1)]) +
  scale_color_manual(values = c("black", tail(blues,1))) +
  scale_x_date(name = NULL, date_breaks="4 month", date_labels = "%b %d %Y", date_minor_breaks = "1 month") +
  ylab("incident deaths") +
  labs(title="Weekly reported COVID-19 deaths in the US: observed and forecasted",
    caption="source: JHU CSSE (observed data), COVID-19 Forecast Hub (forecasts)") +
  theme(legend.position = c(.05,.95), legend.justification = c(0,1), legend.key = element_rect(colour = "transparent", fill = "white"), legend.background = element_rect(alpha("white", 0.5)), legend.box="horizontal")
```

### Hospitalizations

```{r make-US-inc-hospitalization-plot-daily}

inc_hosp_forecast <- dat_hosp %>%
  filter(target %in% inc_hosp_targets)

inc_hosp_truth <- load_truth(
                       truth_source = "HealthData",
                       target_variable = "inc hosp",
                     locations=c("US",all_states))%>%
                    rename(fips = location) %>%
                    left_join(locs, by=c("fips")) %>%
                    filter(fips %in% unique(inc_hosp_forecast$fips), target_end_date <= last_5_saturdays[5])

inc_hosp_all_points <- inc_hosp_truth %>%
  bind_rows(filter(inc_hosp_forecast, class=="point")) %>%
  bind_rows(filter(inc_hosp_truth, target_end_date==last_5_saturdays[5]) %>% mutate(model="COVIDhub-ensemble")) %>%
  mutate(model = relevel(factor(model), ref="Observed Data (HealthData)"))


    
# inc hosp data for code of uncertainty
dummy_inc_hosp <- tibble(
  quantile = quantiles_to_plot, 
  target_end_date=last_5_saturdays[5]) %>%
  right_join(inc_hosp_all_points %>%
      select(-quantile) %>%
      filter(target_end_date == last_5_saturdays[5]))

inc_hosp_quantiles <- inc_hosp_forecast %>%
  dplyr::filter(class=="quantile") %>%
  bind_rows(dummy_inc_hosp) %>%
  dplyr::filter(quantile %in% quantiles_to_plot) %>%
  dplyr::mutate(endpoint_type = ifelse(quantile < 0.5, 'lower', 'upper'),
    alpha = ifelse(endpoint_type == 'lower',
      format(2*quantile, digits=3, nsmall=3),
      format(2*(1-quantile), digits=3, nsmall=3)),
    `Prediction Interval` = fct_rev(paste0((1-as.numeric(alpha))*100, "%"))
  ) %>%
  dplyr::filter(alpha != "1.000") %>%
  dplyr::select(-quantile) %>%
  tidyr::pivot_wider(names_from='endpoint_type', values_from='value')

# daily
ggplot() +
  geom_ribbon(data = inc_hosp_quantiles %>% dplyr::filter(fips=="US"),
    mapping = aes(x = target_end_date,
      ymin=lower, ymax=upper,
      fill=`Prediction Interval`)) +
  geom_line(data=inc_hosp_all_points %>%
      dplyr::filter(fips == "US"),
    mapping = aes(x = target_end_date, y = value, color = model)) +
  geom_point(data=inc_hosp_all_points %>%
      dplyr::filter(fips == "US", !(model=="COVIDhub-ensemble" & target_end_date <= this_monday)),
    mapping = aes(x = target_end_date, y = value, color = model)) +
  scale_fill_manual(values = blues[1:(length(blues)-1)]) +
  scale_color_manual(values = c("black", tail(blues,1))) +
  scale_x_date(name = NULL, date_breaks="4 month", date_labels = "%b %d %Y", date_minor_breaks = "1 month") +
  ylab("incident hospitalizations") +
  labs(title="Daily COVID-19 hospitalization in the US: observed and forecasted",
    caption="source: HealthData (observed data), COVID-19 Forecast Hub (forecasts)") +
  theme(legend.position = c(.05,.95), legend.justification = c(0,1), legend.key = element_rect(colour = "transparent", fill = "white"), legend.background = element_rect(alpha("white", 0.5)), legend.box="horizontal")
```

### Cases

```{r make-US-inc-case-plot}

inc_case_forecast <- dat %>%
  filter(target %in% inc_case_targets)

# get full inc case truth for plotting
inc_case_truth <- load_truth(
                       truth_source = "JHU",
                       target_variable = "inc case",
                     locations=c("US",all_states)) %>%
                    rename(fips = location) %>%
mutate(model = "observed data (JHU)") %>%     #maybe also deleting this line
                    left_join(locs, by=c("fips")) %>%
                    filter(fips %in% unique(inc_case_forecast$fips),target_end_date <= last_5_saturdays[5])


inc_case_all_points <- inc_case_truth %>%
  bind_rows(filter(inc_case_forecast, class=="point")) %>%
  bind_rows(filter(inc_case_truth, target_end_date==last_5_saturdays[5]) %>% mutate(model="COVIDhub-ensemble")) %>%
  mutate(model = relevel(factor(model), ref="observed data (JHU)"))
  
    
# inc case data for code of uncertainty
dummy_inc_case <- tibble(
  quantile = quantiles_to_plot, 
  target_end_date=last_5_saturdays[5]) %>%
  right_join(inc_case_all_points %>%
      select(-quantile) %>%
      filter(target_end_date == last_5_saturdays[5]))

inc_case_quantiles <- inc_case_forecast %>%
  dplyr::filter(class=="quantile") %>%
  bind_rows(dummy_inc_case) %>%
  dplyr::filter(quantile %in% quantiles_to_plot) %>%
  dplyr::mutate(endpoint_type = ifelse(quantile < 0.5, 'lower', 'upper'),
    alpha = ifelse(endpoint_type == 'lower',
      format(2*quantile, digits=3, nsmall=3),
      format(2*(1-quantile), digits=3, nsmall=3)),
    `Prediction Interval` = fct_rev(paste0((1-as.numeric(alpha))*100, "%"))
  ) %>%
  dplyr::filter(alpha != "1.000") %>%
  dplyr::select(-quantile) %>%
  tidyr::pivot_wider(names_from='endpoint_type', values_from='value')

ggplot() +
  geom_ribbon(data = inc_case_quantiles %>% dplyr::filter(fips=="US"),
    mapping = aes(x = target_end_date,
      ymin=as.numeric(lower), ymax=as.numeric(upper),
      fill=`Prediction Interval`)) +
  geom_line(data=inc_case_all_points %>%
      dplyr::filter(fips == "US"),
    mapping = aes(x = target_end_date, y = value, color = model)) +
  geom_point(data=inc_case_all_points %>%
      dplyr::filter(fips == "US", !(model=="COVIDhub-ensemble" & target_end_date <= this_monday)),
    mapping = aes(x = target_end_date, y = value, color = model)) +
  scale_fill_manual(values = blues[1:(length(blues)-1)]) +
  scale_color_manual(values = c("black", tail(blues,1))) +
  scale_x_date(name = NULL, date_breaks="4 month", date_labels = "%b %d %Y", date_minor_breaks = "1 month") +
  ylab("incident cases") +
  labs(title="Weekly reported COVID-19 cases in the US: observed and forecasted",
    caption="source: JHU CSSE (observed data), COVID-19 Forecast Hub (forecasts)") +
  theme(legend.position = c(.05,.95), legend.justification = c(0,1), legend.key = element_rect(colour = "transparent", fill = "white"), legend.background = element_rect(alpha("white", 0.5)), legend.box="horizontal")
```








```{r prep-datatable}

## get last saturday observed cumulative deaths, inc cases, inc hosp

cum_death_start  <- load_truth(
                       truth_source = "JHU",
                       target_variable = "cum death")%>%
                    rename(fips = location) %>%
                    mutate(model = "observed data (JHU)") %>%     #maybe also deleting this line
                    left_join(locs, by=c("fips")) %>%
                    filter(target_end_date == last_5_saturdays[5]) %>%
                    select(target_end_date, location_name.x,fips, value) %>%
                    rename(cum_deaths_at_forecast_start = value,date =target_end_date,location=fips,location_name=location_name.x) 


## get recent observed inc deaths
recent_inc_death_totals <- load_truth(
                              truth_source = "JHU",
                              target_variable = "inc death",
                     locations=c("US",all_states)) %>%
                    rename(fips = location) %>%
                    mutate(last_2wk = target_end_date > last_5_saturdays[3] & target_end_date <= last_5_saturdays[5],
    last_4wk = target_end_date > last_5_saturdays[1] & target_end_date <= last_5_saturdays[5]) %>%    
                      filter(target_end_date >= last_5_saturdays[2]) %>%
                    left_join(locs, by=c("fips")) %>%
   rename(location=fips,location_name=location_name.x) %>%
  select(target_end_date, location,location_name, value, last_2wk, last_4wk) %>%
  group_by(location,location_name) %>%
  summarize(last_2wk_deaths = sum(value*last_2wk),
    last_4wk_deaths = sum(value*last_4wk)) %>%
  ungroup() %>%
    left_join(locs, by=c("location" = "fips")) %>%
  left_join(cum_death_start) %>%
rename(fips = location)

## get recent observed inc cases
recent_inc_case_totals <- load_truth(
                              truth_source = "JHU",
                              target_variable = "inc case",
                     locations=c("US",all_states)) %>%
                    rename(fips = location) %>%
                    mutate(last_1wk = target_end_date > last_5_saturdays[4] & target_end_date <= last_5_saturdays[5],
    last_2wk = target_end_date > last_5_saturdays[3] & target_end_date <= last_5_saturdays[5]) %>%
                      filter(target_end_date >= last_5_saturdays[2]) %>%
                    left_join(locs, by=c("fips")) %>%
   rename(location=fips,location_name=location_name.x) %>%
  select(target_end_date, location,location_name, value, last_1wk, last_2wk) %>%
  group_by(location,location_name) %>%
  summarize(last_1wk_cases = sum(value*last_1wk),
    last_2wk_cases = sum(value*last_2wk)) %>%
  ungroup() %>%
    left_join(locs, by=c("location" = "fips")) %>%
rename(fips = location)

## get recent observed inc hosp
recent_inc_hosp_totals <- load_truth(
                       truth_source = "HealthData",
                       target_variable = "inc hosp",
                     locations=c("US",all_states))%>%
                    rename(fips = location) %>%
                    mutate(last_1wk = target_end_date > last_5_saturdays[4] & target_end_date <= last_5_saturdays[5]) %>%
                      filter(target_end_date >= last_5_saturdays[2]) %>%
                      filter(fips %in% unique(inc_hosp_forecast$fips)) %>%
                    left_join(locs, by=c("fips")) %>%
   rename(location=fips,location_name=location_name.x) %>%
  select(target_end_date, location,location_name, value, last_1wk) %>%
  group_by(location,location_name) %>%
  summarize(last_1wk_hosp = sum(value*last_1wk)/7) %>%
  ungroup() %>%
    left_join(locs, by=c("location" = "fips")) %>%
rename(fips = location)

  
```


```{r process-ensemble-data-deaths}
ensemble_pointdat  <- dat  %>%
  filter(grepl('cum death', target)) %>%
  filter(class=="point") %>%
  select(fips, target, value, timezero)

wide_point_dat <- spread(ensemble_pointdat, target, value) %>%
  left_join(recent_inc_death_totals) %>%
  mutate(next_2wk_deaths = `2 wk ahead cum death` - cum_deaths_at_forecast_start,
    diff_2wk_deaths = next_2wk_deaths - last_2wk_deaths,
    next_4wk_deaths = `4 wk ahead cum death` - cum_deaths_at_forecast_start,
    diff_4wk_deaths = next_4wk_deaths - last_4wk_deaths,
    pop_x_1k = round(Population/1000),
    last_2wk_deaths_rate = round(last_2wk_deaths/Population*100000/14,3),
    last_4wk_deaths_rate = round(last_4wk_deaths/Population*100000/28,3),
    next_2wk_deaths_rate = round(next_2wk_deaths/Population*100000/14,3),
    next_4wk_deaths_rate = round(next_4wk_deaths/Population*100000/28, 3),
    diff_2wk_deaths_rate = round(next_2wk_deaths_rate - last_2wk_deaths_rate, 3),
    diff_4wk_deaths_rate = round(next_4wk_deaths_rate - last_4wk_deaths_rate, 3),
   next_2wk_cum_deaths = `2 wk ahead cum death` - cum_deaths_at_forecast_start) %>%
  select(location_name, Population, pop_x_1k, cum_deaths_at_forecast_start,
    last_2wk_deaths, next_2wk_deaths, diff_2wk_deaths, 
    last_4wk_deaths, next_4wk_deaths, diff_4wk_deaths,
    last_2wk_deaths_rate, next_2wk_deaths_rate,  
    last_4wk_deaths_rate, next_4wk_deaths_rate, 
    diff_2wk_deaths_rate, diff_4wk_deaths_rate, next_2wk_cum_deaths)


#filter quantile data for predicting future weeks
ensemble_quantdat <- dat %>%
  filter(target == "2 wk ahead cum death") %>%
  filter(class == "quantile")

wide_quant_dat <- spread(ensemble_quantdat, target, value) %>%
  left_join(recent_inc_death_totals %>% select(fips, location_name, last_2wk_deaths, cum_deaths_at_forecast_start)) %>% 
  mutate(next_2wk_deaths = `2 wk ahead cum death` - cum_deaths_at_forecast_start)

quant.5_cutoff <- wide_quant_dat %>% 
  filter(quantile == .5, next_2wk_deaths >= last_2wk_deaths) 

quant.5_cutoff_states <- wide_quant_dat %>% 
  filter(quantile == .5, next_2wk_deaths >= last_2wk_deaths, fips < 60) 

quant.5_cutoff_terr <- wide_quant_dat %>% 
  filter(quantile == .5, next_2wk_deaths >= last_2wk_deaths, fips >= 60, fips != 'US') 

quant.25_cutoff <- wide_quant_dat %>% 
  filter(quantile == .25, next_2wk_deaths >= last_2wk_deaths) 

```


```{r process-ensemble-data-cases}

ensemble_pointdat2  <- dat  %>%
  filter(grepl('inc case', target)) %>%
  filter(class=="point") %>%
  select(fips, target, value, timezero)

wide_point_dat2 <- spread(ensemble_pointdat2, target, value) %>%
  left_join(recent_inc_case_totals) %>%
  mutate(next_1wk_cases = `1 wk ahead inc case`,
    pop_x_1k = round(Population/1000),
    last_1wk_cases_rate = round(last_1wk_cases/Population*100000/7,3),
    last_2wk_cases_rate = round(last_2wk_cases/Population*100000/14,3),
    next_1wk_cases_rate = round(next_1wk_cases/Population*100000/7, 3),
    diff_wk_cases_rate = round(next_1wk_cases_rate - last_1wk_cases_rate, 3)) %>%
  select(fips, location_name.x, Population, 
    last_1wk_cases,last_2wk_cases,   next_1wk_cases,  last_1wk_cases_rate,  last_2wk_cases_rate,next_1wk_cases_rate,  diff_wk_cases_rate)


#filter find first quantile where predicting future 1 weeks >= previous week count
ensemble_quantdat2 <- dat %>%
  filter(target == "1 wk ahead inc case") %>%
  filter(class == "quantile")


wide_quant_dat2 <- spread(ensemble_quantdat2, target, value) %>%
  left_join(recent_inc_case_totals %>%
  select(fips,location_name.x, last_1wk_cases)) %>%
  mutate(next_1wk_cases = `1 wk ahead inc case`)

quant_cutoff2 <- wide_quant_dat2 %>%
  filter(next_1wk_cases >= last_1wk_cases) %>%
  mutate(prob_case_up = 1-quantile) %>%
select( quantile, location_name.x, prob_case_up,class)

quant_cutoff2min<-quant_cutoff2[!duplicated(quant_cutoff2$location_name.x),]

wide_point_dat2p <- merge(wide_point_dat2, quant_cutoff2min, by="location_name.x")

```

```{r process-ensemble-data-hosp} 

ensemble_pointdat_hosp <- dat_hosp  %>%
  filter(grepl('inc hosp', target)) %>%
  filter(class=="point") %>%
  
                    mutate(next_2wk = target_end_date > next_saturday & target_end_date <= next_2saturday) %>%
                      filter(target_end_date > next_saturday) %>%
                      filter(target_end_date <= next_2saturday) %>%
                      filter(fips %in% unique(inc_hosp_forecast$fips)) %>%
  select(fips, target, value, timezero,next_2wk)  %>%
                    left_join(locs, by=c("fips")) %>%
  group_by(fips) %>%
  summarize(next_2wk_hosp = sum(value*next_2wk)/7) %>%
  ungroup() 

`%!in%` <- Negate(`%in%`)
wide_point_dat_hosp <- ensemble_pointdat_hosp %>%
  left_join(recent_inc_hosp_totals, by=c("fips")) %>%
  mutate(pop_x_1k = round(Population/1000),
    last_1wk_hosp_rate = round(last_1wk_hosp/Population*100000,3),
    next_2wk_hosp_rate = round(next_2wk_hosp/Population*100000,3),
    diff_wk_hosp_rate = round(next_2wk_hosp_rate - last_1wk_hosp_rate, 3)) %>%
   filter(fips %!in% c("66","69"))  %>% #filter out Guam and North Mariana Islands
  select(fips, location_name.x, Population, 
    last_1wk_hosp, next_2wk_hosp, last_1wk_hosp_rate,   next_2wk_hosp_rate, diff_wk_hosp_rate)


# #filter find first quantile where predicting future 2 weeks >= previous week count
# ensemble_quantdat_hosp <- dat_hosp %>%
#   filter(target_end_date > next_saturday) %>%
#   filter(target_end_date <= next_2saturday) %>%
#  filter(fips %in% unique(inc_hosp_forecast$fips)) %>%
#   filter(class == "quantile")
#   
# 
# wide_quant_dat_hosp <- spread(ensemble_quantdat_hosp, target, value) %>%
#   left_join(recent_inc_hosp_totals %>% 
#   select(fips,location_name.x, last_1wk_hosp)) %>%
#   mutate(next_2wk_hosp = `2 wk ahead inc hosp`)
# 
# quant_cutoff_hosp <- wide_quant_dat_hosp %>%
#   filter(next_2wk_hosp >= last_1wk_hosp) %>%
#   mutate(prob_hosp_up = 1-quantile) %>%
# select( quantile, location_name.x, prob_hosp_up,class)
# 
# quant_cutoff_hosp_min<-quant_cutoff_hosp[!duplicated(quant_cutoff_hosp$location_name.x),]
# 
# wide_point_dat_hosp_p <- merge(wide_point_dat_hosp, quant_cutoff2min, by="location_name.x")

```


## State level {.tabset .tabset-fade}

The ensemble model estimates that `r nrow(quant.5_cutoff_states)` out of 50 states and `r nrow(quant.5_cutoff_terr)` out of 7 territories of the US have a greater than 50% chance of having more deaths in the next two weeks compared to the past two weeks.The model forecasts that `r nrow(quant.25_cutoff)` states and territories have a greater than 75% chance of an increase over the next two weeks (`r paste(quant.25_cutoff$location_name, collapse=", ")`). 

The sortable and searchable table below shows the total number of reported COVID-19 deaths at the US level and by state over the last two weeks (ending Saturday, `r format(last_5_saturdays[5], "%B %d, %Y")`) and the forecasted counts for the subsequent two weeks (ending `r format(last_5_saturdays[5]+14, "%B %d, %Y")`).

```{r make-datatable-inc-death-counts}
death_max_2wk <- max(abs(wide_point_dat$diff_2wk_deaths))

brks <- seq(-death_max_2wk, death_max_2wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
#clrs <- scico(n=length(brks)+1, palette="roma")
clrs <- colorRampPalette(colors = rev(RColorBrewer::brewer.pal(n=3, "RdBu")))(length(brks)+1)

table1_cap <- paste0("Table 1: US national and state-level observed deaths to date and for the previous two weeks (ending ", format(last_5_saturdays[5], "%B %d, %Y") ,") and the next two weeks (ending ", format(last_5_saturdays[5]+14, "%B %d, %Y"), ").")

datatable(wide_point_dat %>% 
    select(location_name, Population,
      cum_deaths_at_forecast_start, 
      last_2wk_deaths, next_2wk_deaths, diff_2wk_deaths) %>%
    arrange(desc(diff_2wk_deaths)),
  caption = table1_cap,
  options = list(
    autoWidth = TRUE,
    columnDefs = list(list(width = '100px', targets = c(0, 1, 2, 3, 4, 5)))
  ),  #width=paste0(c(10, 100, 100, 100), 'px'),
  rownames=FALSE,
  colnames=c('state'='location_name', 
    #'Population, \'000'='pop_x_1k',
    'Population' = 'Population',
    'Total COVID-19 deaths'='cum_deaths_at_forecast_start',
    'COVID-19 deaths, last 2 weeks'='last_2wk_deaths',
    'COVID-19 deaths, next 2 weeks'='next_2wk_deaths',
    'Difference' = 'diff_2wk_deaths')) %>%
##  formatStyle("Daily deaths, last 2 weeks", backgroundColor = styleInterval(brks, clrs)) %>%
##  formatStyle("Daily deaths, next 2 weeks", backgroundColor = styleInterval(brks, clrs))  %>%
  formatStyle("Difference", backgroundColor = styleInterval(brks, clrs)) %>%
  formatCurrency('Total COVID-19 deaths',currency = "", interval = 3, mark = ",", digits=0) %>%
  formatCurrency('Population',currency = "", interval = 3, mark = ",", digits=0) %>%
  formatCurrency('COVID-19 deaths, last 2 weeks',currency = "", interval = 3, mark = ",", digits=0) %>%
  formatCurrency('COVID-19 deaths, next 2 weeks',currency = "", interval = 3, mark = ",", digits=0)

```

Looking at the rates allows for easier comparison across states, where you can see which states have had or are predicted to have proportionally higher rates in comparison to other states. 

When interpreting probability of an increase, it's important to note that the increase or decrease is relative to the location-specific recent observed incidence, which varies across locations.

### Death rates

The sortable and searchable death table below shows the total number of reported COVID-19 deaths at the US level and by state as of Saturday, `r format(last_5_saturdays[5], "%B %d, %Y")` ("Total COVID-19 Deaths") as well as the rate of reported COVID-19 deaths in the population (standardized per 100,000 population) over the last two weeks and over the next two weeks. 
```{r make-datatable-inc-death-rates}

## color for rates
death_rate_max_2wk <- max(c(wide_point_dat$last_2wk_deaths_rate, wide_point_dat$next_2wk_deaths_rate))
brks <- seq(0, death_rate_max_2wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

## colors for rate difference
death_rate_diff_2wk <- max(abs(wide_point_dat$diff_2wk_deaths_rate))
brks1 <- seq(-death_rate_diff_2wk, death_rate_diff_2wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs1 <- colorRampPalette(colors = rev(RColorBrewer::brewer.pal(n=3, "RdBu")))(length(brks1)+1)


table2a_cap <- paste0("Table 2a: US national and state-level observed and predicted daily death rates for the previous two weeks (ending ", format(last_5_saturdays[5], "%B %d, %Y") ,") and the next two weeks (ending ", format(last_5_saturdays[5]+14, "%B %d, %Y"), ").")

datatable(wide_point_dat %>% 
    select(location_name, Population,
      cum_deaths_at_forecast_start, 
      last_2wk_deaths_rate, next_2wk_deaths_rate, diff_2wk_deaths_rate) %>%
    arrange(desc(diff_2wk_deaths_rate)),
  caption = table2a_cap,
  options = list(
    autoWidth = TRUE,
    columnDefs = list(list(width = '100px', targets = c(0, 1, 2, 3, 4, 5)))
  ),  #width=paste0(c(10, 100, 100, 100), 'px'),
  rownames=FALSE,
  colnames=c('state'='location_name', 
    #'Population, \'000'='pop_x_1k',
    'Population' = 'Population',
    'Total COVID-19 deaths'='cum_deaths_at_forecast_start',
    'Daily deaths per 100k, last 2 weeks'='last_2wk_deaths_rate',
    'Daily deaths per 100k, next 2 weeks'='next_2wk_deaths_rate',
    'Death rate difference' = 'diff_2wk_deaths_rate')) %>%
  ## formatStyle("Daily deaths per 100k, last 2 weeks", backgroundColor = styleInterval(brks, clrs)) %>%
  ## formatStyle("Daily deaths per 100k, next 2 weeks", backgroundColor = styleInterval(brks, clrs))  %>%
  formatStyle('Death rate difference', backgroundColor = styleInterval(brks1, clrs1))  %>%
  formatCurrency('Total COVID-19 deaths',currency = "", interval = 3, mark = ",", digits=0) %>%
  formatCurrency('Population',currency = "", interval = 3, mark = ",", digits=0)
```


### Hospitalization rates

The sortable and searchable hospitalization table below shows the 7 day average hospitalization count and rate of reported COVID-19 hospitalization in the population (standardized per 100,000 population) over the last week (`r format(last_5_saturdays[4]+1, "%B %d, %Y")` to `r format(last_5_saturdays[5], "%B %d, %Y")`) and two weeks ahead (`r format(last_5_saturdays[5]+8, "%B %d, %Y")` to `r format(last_5_saturdays[5]+14, "%B %d, %Y")`).
```{r make-datatable-inc-hosp-rates}

## color for rates
hosp_rate_max_1wk <- max(c(wide_point_dat_hosp$last_1wk_hosp_rate, wide_point_dat_hosp$next_2wk_hosp_rate))
brks_hosp <- seq(0, hosp_rate_max_1wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs_hosp <- round(seq(255, 40, length.out = length(brks_hosp) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

## colors for rate difference
hosp_rate_diff_2wk <- max(abs(wide_point_dat_hosp$diff_wk_hosp_rate))
brks1_hosp <- seq(-hosp_rate_diff_2wk, hosp_rate_diff_2wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs1_hosp <- colorRampPalette(colors = rev(RColorBrewer::brewer.pal(n=3, "RdBu")))(length(brks1_hosp)+1)

table2c_cap <- paste0("Table 2c: US national and state observed and predicted 7 day average COVID-19 hospitailzation for the previous week and two weeks ahead.")

table2c_Observed <- paste0("Observed: (", format(last_5_saturdays[4]+1, "%m-%d-%Y") ," to ",format(last_5_saturdays[5], "%m-%d-%Y"),")")

table2c_Predicted <- paste0("Predicted: (", format(last_5_saturdays[5]+8, "%m-%d-%Y") ," to ",format(last_5_saturdays[5]+14, "%m-%d-%Y"),")")

# a custom table container
sketch_tab3 = htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(rowspan = 2, "Geography"),
      th(rowspan = 2, "Population"),
      th(colspan = 2, table2c_Observed),
      th(colspan = 2, table2c_Predicted),
      th(rowspan=2,"Hospitalization rate difference")
    ),
    tr(
 lapply((c("Daily count (Obs)","Daily hospitalization per 100K (Obs)","Daily count (Pred)","Daily hospitalization per 100K (Pred)")), th)))))

datatable(wide_point_dat_hosp %>%
    select(location_name.x, Population,
    last_1wk_hosp,last_1wk_hosp_rate, next_2wk_hosp, next_2wk_hosp_rate, diff_wk_hosp_rate) %>%
    arrange(desc(next_2wk_hosp_rate)),
  caption = table2c_cap,
  options = list(
    autoWidth = TRUE,
    columnDefs = list(list(width = '100px', targets = c(0, 1, 2, 3, 4, 5,6)))),

  #width=paste0(c(10, 100, 100, 100), 'px'),
  rownames=FALSE,
  colnames=c('Geography'='location_name.x',
    #'Population, \'000'='pop_x_1k',
    'Daily hospitalization count (Obs)'='last_1wk_hosp',
    'Rate per 100K (Obs)'='last_1wk_hosp_rate',
    'Daily hospitalization count (Pred)'='next_2wk_hosp',
    'Rate per 100K (Pred)'='next_2wk_hosp_rate',
    'Hospitalization rate difference'='diff_wk_hosp_rate') ,
     container=sketch_tab3) %>%
  ## formatStyle("Daily deaths per 100k, last 2 weeks", backgroundColor = styleInterval(brks, clrs)) %>%
  ## formatStyle("Daily deaths per 100k, next 2 weeks", backgroundColor = styleInterval(brks, clrs))  %>%
  formatStyle('Hospitalization rate difference', backgroundColor = styleInterval(brks1_hosp, clrs1_hosp))  %>%
  formatCurrency('Population',currency = "", interval = 3, mark = ",", digits=0)%>%
  formatCurrency('Daily hospitalization count (Obs)',currency = "", interval = 3, mark = ",", digits=0)%>%
  formatCurrency('Rate per 100K (Obs)',currency = "", interval = 3, mark = ",", digits=3)%>%
  formatCurrency('Rate per 100K (Pred)',currency = "", interval = 3, mark = ",", digits=3)%>%
 formatCurrency('Daily hospitalization count (Pred)',currency = "", interval = 3, mark = ",", digits=0)
```

### Case rates

The sortable and searchable case table below shows the weekly case count and rate of reported COVID-19 case in the population (standardized per 100,000 population) over the last week (`r format(last_5_saturdays[4]+1, "%B %d, %Y")` to `r format(last_5_saturdays[5], "%B %d, %Y")`) and one week ahead (`r format(last_5_saturdays[5]+1, "%B %d, %Y")` to `r format(last_5_saturdays[5]+7, "%B %d, %Y")`).
```{r make-datatable-inc-case-rates}

## color for rates
case_rate_max_1wk <- max(c(wide_point_dat2$last_1wk_cases_rate))
brks_case <- seq(0, case_rate_max_1wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs_case <- round(seq(255, 40, length.out = length(brks_case) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

## colors for rate difference
case_rate_diff_wk <- max(abs(wide_point_dat2$diff_wk_cases_rate))
brks1_case <- seq(-case_rate_diff_wk, case_rate_diff_wk, length.out = 100) #quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
clrs1_case <- colorRampPalette(colors = rev(RColorBrewer::brewer.pal(n=3, "RdBu")))(length(brks1_case)+1)

table2b_cap <- paste0("Table 2b: US national and state observed and predicted COVID-19 cases for the previous week and one week ahead.")

table2b_Observed <- paste0("Observed: (", format(last_5_saturdays[4]+1, "%m-%d-%Y") ," to ",format(last_5_saturdays[5], "%m-%d-%Y"),")")

table2b_Predicted <- paste0("Predicted: (", format(last_5_saturdays[5]+1, "%m-%d-%Y") ," to ",format(last_5_saturdays[5]+7, "%m-%d-%Y"),")")

# a custom table container
sketch_tab2 = htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(rowspan = 2, "Geography"),
      th(rowspan = 2, "Population"),
      th(colspan = 2, table2b_Observed),
      th(colspan = 2, table2b_Predicted),
      th(rowspan=2,"Case rate difference")
    ),
    tr(
 lapply((c("Weekly case count (Obs)","Daily cases per 100K (Obs)","Weekly case count (Pred)","Daily cases per 100K (Pred)")), th)))))

datatable(wide_point_dat2p %>%
    select(location_name.x, Population,
    last_1wk_cases,
      last_1wk_cases_rate, next_1wk_cases, next_1wk_cases_rate, diff_wk_cases_rate) %>%
    arrange(desc(next_1wk_cases_rate)),
  caption = table2b_cap,
  options = list(
    autoWidth = TRUE,
    columnDefs = list(list(width = '100px', targets = c(0, 1, 2, 3, 4, 5,6)))),

  #width=paste0(c(10, 100, 100, 100), 'px'),
  rownames=FALSE,
  colnames=c('Geography'='location_name.x',
    #'Population, \'000'='pop_x_1k',
    'Weekly case count (Obs)'='last_1wk_cases',
    'Daily cases per 100K (Obs)'='last_1wk_cases_rate',
    'Weekly case count (Pred)'='next_1wk_cases',
    'Daily cases per 100K (Pred)'='next_1wk_cases_rate',
    'Case rate difference'='diff_wk_cases_rate') ,
     container=sketch_tab2) %>%
  ## formatStyle("Daily deaths per 100k, last 2 weeks", backgroundColor = styleInterval(brks, clrs)) %>%
  ## formatStyle("Daily deaths per 100k, next 2 weeks", backgroundColor = styleInterval(brks, clrs))  %>%
  formatStyle('Case rate difference', backgroundColor = styleInterval(brks1_case, clrs1_case))  %>%
  formatCurrency('Population',currency = "", interval = 3, mark = ",", digits=0)%>%
  formatCurrency('Weekly case count (Obs)',currency = "", interval = 3, mark = ",", digits=0)%>%
  formatCurrency('Daily cases per 100K (Obs)',currency = "", interval = 3, mark = ",", digits=3)%>%
  formatCurrency('Daily cases per 100K (Pred)',currency = "", interval = 3, mark = ",", digits=3)%>%
 formatCurrency('Weekly case count (Pred)',currency = "", interval = 3, mark = ",", digits=0)
```

# Methods & Acknowledgement

This report was reproducibly and dynamically generated using RMarkdown. The code for the report can be found [here](https://github.com/reichlab/covid19-forecast-hub/tree/master/code/reports).


```{r}
# htmltools::includeScript("statcounter.js")
```

